---
title: "Classification Models"
output: html_notebook
---

```{r}
library(tidyverse)
library(caret)
library(AppliedPredictiveModeling)
library(pROC)
```

```{r}
simulated_train <- quadBoundaryFunc(500)
simulated_test <- quadBoundaryFunc(1000)

head(simulated_train)
```

```{r}
ranf_model <- train(simulated_train[,1:2], simulated_train$class,
                   method = "rf",
                   trControl = trainControl(method = "cv", number = 10),
                   tuneLength = 9)

ranf_model
```

```{r}
confusionMatrix(ranf_model)
```
I need a way to show the ROC curve.

```{r}
roc_curve <- roc(response = simulated_test$class,
                 predictor = simulated_test$prob,
                 levels = rev(levels(simulated_test$class)))

plot.roc(roc_curve)
```
This can be used quite nicely to explain class probabilities.

```{r}
roc_curve
```
AUC is 0.94 which is quite high. The higher the AUC is, the better the model. I assume that when I'm deploying the model, I would use
a threshold that maximises the tradeoff between sensitivity and specificity.

There is a problem with this. What if there is a curve that has a lower area but has a point where its threshold is better than the higher AUC. 

ROC curves are used to maximise the trade-off between sensitivity and specificity. If you increase sensitivity, specificity will be reduced. 
If you increase sensitivity, a higher proportion of identified positives will be real positives. On the other hand, you get less true negatives, more false negatives. I'm way too tired for this.

Alright, how do I select the best threshold given a single ROC curve?

## Training models

You can specify ROC to be the evaluation criteria for the models, not the accuracy. However, even after that, the probability cutoff should be selected.
How do you do that? Is it automatically selected? Do I need to select it myself?

You can also just tune your model for accuracy. BUT. What if you can also get much better results by additionally shifting class probability thresholds.
You can do this by manually shifting the threshold. But this is post-cross validation. Therefore, wouldn't it potentially lead to overfitting the model?

Ok, I read ahead a few chapters. The author uses a different holdout set of data to deal with class imbalance called the evaluation set. This comes after model training but before model testing. This is a tricky situation. In this case even though you cross-validated your model, due to the evaluation tuning, you have to test it on a separate test dataset. Interesting. I guess this is the price of dealing with data with extreme class imbalance.

The class-probability shifts definitely need to be used when the cost of a false positive is not the same as the cost of a false negative.
It could just be reclassified manually. I will dig into this a bit deeper.

# Nonlinear Classification Models





































