---
title: "Class-exercises"
output: html_notebook
---

```{r}
library(tidyverse)
library(AppliedPredictiveModeling)
library(caret)
```

# 12.1

```{r}
data(hepatic)
```

## A

There are 3 levels of the response variable, therefore, this is a 3 class prediction issue.

```{r}
table(injury)
```
There is quite an imbalance. The severe class is not highly represented.

What I would use for slicing up the data into training and test sets are stratified random sampling.

## B - Which classification statistic would you optimise?

Things I could optimisise:

* AUC - It works for 2 class models only.
* Accuracy
* I'm guessing the kappa statistic could also be optimised for

Since I can't optimise for AUC, I would optimise for plain old accuracy.

Building models:

## Correlation plots

```{r}
corrplot::corrplot(cor(bio), method = "color")
```

There are some empty rows going on that should be removed.

```{r}
bio <- bio[,-nearZeroVar(bio)]
```
82 variables were removed

```{r}
findCorrelation(cor(bio), cutoff = 0.85) %>% length()
```
10 - 20 variables are highly correlated. Let's be stringent with removing variables.


