---
title: "Machine Learning with R and Python"
date: "February, 2019"
author: "Naoki Ohno"
output: html_notebook
---

## Loading data

```{r include=FALSE}
library(tidyverse)
library(reticulate)

#use_python("C:/Users/naoki.ohno/AppData/Local/Continuum/anaconda3/envs/py37/python.exe")
use_condaenv("py37")
py_config()
```

```{r}
## Data preparation

# Reading in the data
test_init <- read_csv("data/test.csv")
train_init <- read_csv("data/train.csv")

test_init$Survived <- NA
data <- rbind(train_init, test_init)

# Order columns 
data <- data %>% select(-Survived, Survived)
data <- as.data.frame(data)
```

## Feature engineering

```{r}
data <- data %>% mutate(no_family = SibSp + Parch)
data <- data %>% select(-c(PassengerId, SibSp, Parch, Ticket, Cabin))

data$Pclass <- factor(data$Pclass)
data$Name <- factor(data$Name)
data$Sex <- factor(data$Sex)
data$Embarked <- factor(data$Embarked)
data$Survived <- factor(data$Survived)
```

```{r}
data %>% filter(Pclass == 3, Embarked == "S") %>% drop_na(Fare) %>% summarise(median_fare = median(Fare))
data[is.na(data$Fare), 'Fare'] <- 8.05
```

```{r}
data$Name <- as.character(data$Name)
data$title <- sub("^.*, ", "", data$Name)
data$title <- sub("\\..*", "", data$title)

data$title <- gsub("Mlle", "Miss", data$title)
data$title <- gsub("Mme", "Mrs", data$title)
data$title <- gsub("Ms", "Miss", data$title)

# Aggregating important people
data$title <- gsub("Capt|Col|Don|Dona|Dr|Jonkheer|Lady|Major|Rev|Sir|the Countess", "High_rank", data$title)

data[is.na(data$Embarked), "Embarked"] <- "C"

## Imputing missing age
age_lm <- lm(Age ~ Sex + Fare + no_family + title, data = na.omit(data))
data[is.na(data$Age), "Age"] <- predict(age_lm, data[is.na(data$Age),])

# If age is negative, turn into 0
data$Age <- ifelse(data$Age < 0, 0, data$Age)

# Round age to half number
data$Age <- round(data$Age, 1)

data <- data %>% select(Survived, Sex, Age, Fare, Embarked, no_family, title)

train <- data[1:nrow(train_init),]
```

The dataset is now clean and can be used for machine learning. The feature engineering is not yet ideal but it's good enough to run initial models using Python's Scikit-learn package.

```{python}
#train = r.train

import numpy as np
import pandas as pd
from sklearn.ensemble import RandomForestClassifier


```

```{python}

```

Ok, I think a different approach might work better. I will follow this book instead of going off into my own tangent.

You can either learn machine learning with R or Pandas and Numpy.
It would be very hard to start with Scikit learn using Python.





