---
title: "Data Pre-Processing Exercises"
output: html_notebook
---

```{r}
library(caret)
library(tidyverse)
library(AppliedPredictiveModeling)
library(mlbench)
library(corrplot)
```

```{r}
data(Glass)
data <- Glass

data$Type <- as.numeric(as.character(data$Type))
corrplot(cor(data), method = "number")
```

We have 4 seemingly important predictors:

* Mg 
* Al
* Na
* Ba


Relationships between predictors:

* Ca and Refractive index - Possibly, the added Calcium makes it cloudy as CaCo3 is very opaque.

```{r}
pairs(data)
```

The response variable glass type is a categorical variable, therefore linear regression is not applicable.

```{r}
GGally::ggpairs(data, progress = F)
```

These initial plots are super important.

Skewed variables:

* Fe
* Ba 
* K
* Mg is bimodal
* RI

Outliers: Negative values possibly, nothing to note.

There is a linear relationship between type and Si as well as Al.

I guess I could transform the variables with Box-Cox transformation but it won't really do anything for tree-based models.

I don't get it. Doesn't the box cox transformation mess the original data up? How do we know?
